---
title: "data manipulation with R 2nd"
author: "qiufei"
date: "September 10, 2015"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    theme: readable
    toc: yes
  word_document: default
---

R is a domain-specific programming language, specially designed to perform statistical analysis on data.



# introduction to R data types and basic operation

R has a very steep learning curve like python,octave,and sas. However unlike octave and sas, we can find a larger number of freely available resources and tutorials on the web to get help.

a collection of data values stored in a single object is known as vector.

## data type

whatever we do in R, is stored as objects. An R objects is anything thant can be assigned to a variable of interest.

the mode and class are the most important attributes of an R objects.

if an R object contains both numeric and logical elements, the mode of that object will be numberic and ,in this case, the logical element automatically gets converted into a numeric element. the logical element `TRUE` convets to 1 and `FALSE` converts to 0.

on the other hand, if any R object contains a character element, along with both numerica and logical elements, it automatically converts to character mode.

```{r}
xz=c(1,3,TRUE,5,FALSE,9)
mode(xz)

xw=c(1,2,TRUE,FALSE,'a')
mode(xw)
```

## an interesting calculation

```{r}
1-5*0.2
# you think they are same
1-0.2-0.2-0.2-0.2-0.2
```

## a logical vector can be thought as a vector of an index

```{r}
age=c(20,30,22,25)
willing=c(TRUE,FALSE,TRUE,FALSE)
age[willing]
```

## as.numeric() function only returns internal values of the factor

```{r}
num.factor=factor(c(5,7,9,5,6,7,3,5,3,9,7))
num.factor

as.numeric(num.factor)

levels(num.factor)

as.character(num.factor)

# now to convert the 'num.factor' to numeric there are two methods
# method 1
as.numeric(as.character(num.factor))

# method 2
as.numeric(levels(num.factor)[num.factor])

# explain of method 2
levels(num.factor)

levels(num.factor)[num.factor]

```
## missing values in R

In R, a numeric missing value is represented by `NA`, while character missing values are represented by `<NA>`. 

# basic data manipulation

## remove unused factors

```{r}
char.obj="datamanipulation"

substring(char.obj,1:nchar(char.obj),1:nchar(char.obj))

# some explain
substring("abcdef", 1:6, 1:6)

substring("abcdef", 1:3, 1:6)

factor.obj=factor(substring(char.obj,1:nchar(char.obj),1:nchar(char.obj)),levels=letters)

table(factor.obj)

factor.obj1=factor(factor.obj)

factor.obj2=droplevels(factor.obj)

table(factor.obj1)
table(factor.obj2)
```

## lubridate package

```{r}
library(lubridate)
mdy("1-1-1970")
mdy("01-01-1970")
mdy("1-01-1970")
mdy("1-01-70")


# change month

date=dmy("23-07-2013")
date

month(date)
month(date)=8
date

```

## keeping only values greater than 6

usually, numerical integers are used for subscripting, but logical vectors can also be used for the same purpose.

```{r}
num10=c(3,2,5,3,9,6,7,9,2,3)
num10>6
num10[num10>6]
```

# chapter 3 applying the split-apply-combine strategy


## split-apply-combine using base function of R
```{r}
library(xlsx)
price=read.xlsx(file='chapter3.xls',sheetIndex = 1)

### step 1: split

# detailed code to implement the split-apply-combine approach

shanghai=subset(price,region=='shanghai',select = equipment:agriculture)

zhejiang=subset(price,region=='zhejiang',select = equipment:agriculture)

jiangsu=subset(price,region=='jiangsu',select = equipment:agriculture)

# less code to implement the split-apply-combine approach

price.split=split(price,price$region)
    # if region is not factor ,you should use as.factor(price$region)

### step 2 : apply
# apply mean funtion to calculate mean

price.apply=lapply(price.split,function(x) colMeans(x[,3:6]))

### step 3 : combine

price.combine=do.call(rbind,price.apply)

price.combine

str(price.combine)

```


## split-apply-combine using `plyr` package

the `plyr` package works on every type of data structure, whereas the `dplyr` package is designed to workonly on data frames.


the most important utility of the `plyr` package is that a single line of code can perform all the split,apply and combine steps.

```{r}
library(plyr)
price.mean=ddply(price,.(region),function(x) colMeans(x[,3:6]))

price.mean

str(price.mean)
```

## an array example

```{r}

class(iris3)
dim(iris3)


iris.mean1=adply(iris3,.margins = 3,colMeans)
iris.mean1
class(iris.mean1)

iris.mean2=aaply(iris3,.margins = 3,colMeans)
iris.mean2
class(iris.mean2)  # note that here the class is showing "matrix",since the output is a two dimensional array which represents matrix. 

iris.mean3=alply(iris3,.margins = 3,colMeans)
iris.mean3
class(iris.mean2)


```


## three main arguments for the `plyr` common functions 

1. a*ply(.data, .margins, .fun, ..., .progress = "none")
2. d*ply(.data, .variables, .fun, ..., .progress = "none")
3. l*ply(.data, .fun, ..., .progress = "none")

我的理解:list最本质的特征是它能处理非齐整的数据,所以没办法按照margin或者variable来分层.或者这么说,如果是一个完全齐整的list的,何不转化为array或datafame来处理?

If we want to monitor the progress of the processing task, the `.progress` argument should be specified. It will not show the progress status by default,that is `.progress = "none"'`.



